# Data

## Description
1. International Renewable Energy Agency (IRENA)
Link: https://www.irena.org/Data/Downloads/IRENASTAT (link to IRENA data query tool)

The data is collected by the international renewable energy agency IRENA from its member countries. IRENA is an intergovernmental organization that exists to support countries with their transition to using sustainable energy. There are currently 169 member countries and the EU. IRENA is able to collect their data through their questionnaire for these member countries. In addition, desk research is also conducted to complement the questionnaire to provide a comprehensive report. IRENA updates its database annually. 

There is a data query on the IRENA website for us to look at data related to heat generation or electricity generation in general. There is also data that looks at sustainable energy as a percentage of overall energy generation of all the member countries. We are able to select these data through a data query on the IRENA website. The data query allows us to download data directly in excel or csv format. To import the data, we will download the data as a csv file and load it in to R with read.csv() function. The data is in wide format that is sorted by countries, then by different groups of indicators (% wind generation, etc.), and finally by yearly data (this is consistent across the data we found using the data query). It has data starting from the 2000s to 2023 (previous year).

2. Global Carbon Project (GCP)
Link: https://www.icos-cp.eu/science-and-impact/global-carbon-budget/2022

The data is collected by the Global Carbon Project. The GCP is an international research project that aims to understand patterns and interactions connected to Global Carbon emissions. They are able to collect their data by working with different institutions and consolidating their observations. The data should be updated yearly to the previous year we are currently on eg. (2023 as we are in 2024 right now). However, this data is not available yet, so it might be a potential problem when we perform our data analysis. In addition, there are also a few countries that doesn't have any data which we will have to deal with.

We are able to download data directly on the global carbon budget website. The data contains information on global carbon emissions data as well as emissions of fossil fuels globally. The website allows us to download the information as an excel sheet. We can use the read.xlsx() function to read in and process these excel files into R. The data is sorted by years with data information situated on top of the actual data. There is also territorial emissions data by country. There are also multiple sheets within an excel file we decided to use only one in each excel file as it is most relevant to our analysis. 


## Missing value analysis

### Global Carbon Project Data 
```{r}
library(readxl)
library(redav)
```

```{r}
df_globalcarbon <- read_excel("./raw_data/Global_Carbon_Budget_2022v1.0.xlsx",
                              sheet = 2,
                              skip = 20)
```
```{r}
#| fig-width: 10
#| fig-height: 10
plot_missing(df_globalcarbon, percent = FALSE)
```

The Global Carbon Data denotes the overall fossil fuel emissions globally. There are no missing values as shown colSums of NA values (0 missing values in each of the columns).

```{r}
df_territorialemissions <- read_excel("./raw_data/National_Fossil_Carbon_Emissions_2022v1.0.xlsx",
                              sheet = 2,
                              skip = 11)
```


```{r}
#| fig-width: 10
#| fig-height: 10
plot_missing(df_territorialemissions, percent = FALSE)
```
There are quite a lot of missing values. As these emission data is by year, we tried to filter by year to see how much data is left. The number of years that are missing in data is different for each country. 


### Electricity statistics (MW/GWh) by Region, Technology, Data Type and Year

```{r}
#| message: false
# load libraries
library(tidyverse)
library(data.table)

# read raw file
df_electstats_region <- read_csv(
  "./raw_data/irena_electstats_region.csv"
)

attr(df_electstats_region, 'spec') <- NULL
```


```{r}
# electricity statistics column is saved as character. 
# in particular, some rows have value equal "-", we label them as NA 
# convert electricity statistics to numeric
df_electstats_region <- df_electstats_region |> 
  mutate(
    across(
      starts_with('Elec'),
      ~ifelse(.x=='-',NA,.x)
    )
  ) |> 
  mutate(
    across(
      starts_with('Elec'),
      as.numeric
    )
  ) 
```

Let's review the proportion of NAs by column:

```{r}
# review proportion of NAs per column: global
df_electstats_region|> 
  summarise(
    across(
      everything(),
      ~sum(ifelse(is.na(.x),1,0))
    ),
    n=n()
  ) |> 
  pivot_longer(
    cols=-n,
    names_to = 'variable',
    values_to = 'n_na'
  ) |> 
  mutate(prop_na=n_na/n) |> 
  data.table()
```
Only Electricity statistics column presents NAs, less than 11% of the total observations.

Let's review the pattern by Year

```{r}
# review proportion of NAs by Year
df_electstats_region_year <- df_electstats_region |> 
  group_by(Year) |> 
  summarise(
    n = n(),
    n_country = n_distinct(`Region`),
    across(
      starts_with('Elect'),
      ~sum(ifelse(is.na(.x),1,0))
    )
  ) |>
  rename(n_na=starts_with('Elect')) |> 
  mutate(prop_na=n_na/n)

# graph share of NAs in Electric Statistics per Year
df_electstats_region_year |> 
  ggplot(aes(x=Year,y=prop_na)) +
  geom_line() +
  theme_bw() +
  ggtitle('% NAs in Electricity Statistics per Year') +
  scale_y_continuous('% NAs',labels=scales::percent_format())
```

In general, we observe a downtrend across years but year 2023 where there's a big jump.

Let's assess whether % of NAs depend on which Technology is measured

```{r}
#| fig-height: 7
# review proportion of NAs by Year and Technology
df_electstats_region_year_tec <- df_electstats_region |> 
  group_by(Year,Technology) |> 
  summarise(
    n = n(),
    across(
      starts_with('Elect'),
      ~sum(ifelse(is.na(.x),1,0))
    )
  ) |>
  rename(n_na=starts_with('Elect')) |> 
  mutate(prop_na=n_na/n)

#  % share of NAs per by Year and Technology
df_electstats_region_year_tec |> 
  mutate(Year=factor(Year)) |> 
  ggplot(aes(x=Technology,y=Year,fill=prop_na)) +
  geom_tile() +
  scale_fill_gradient2(
    "% of NAs",
    low='green',
    mid='white',
    high='orange',
    midpoint = 0.5,
    labels=scales::percent_format()
  ) +
  theme_bw() +
  ggtitle('% NAs in Electricity Statistics\nper Year and Technology') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Most Technologies have 0% of NAs.Geothermal, Other non-renewable and Pumped Store have around 15% of NAs. Marine and Nuclear has the most share of NAs with over 25%.

Let's analyze regions, could some regions more prone to missing values?

```{r}
#| fig-width: 7
# review proportion of NAs by Year and Region
df_electstats_region_year_r <- df_electstats_region |> 
  group_by(Year,Region) |> 
  summarise(
    n = n(),
    across(
      starts_with('Elect'),
      ~sum(ifelse(is.na(.x),1,0))
    )
  ) |>
  rename(n_na=starts_with('Elect')) |> 
  mutate(prop_na=n_na/n) |> 
  mutate(Region = case_when(
    str_detect(Region,'Central') ~ 'Central America',
    TRUE ~ Region
  ))

# graph % NA per Year and Country
df_electstats_region_year_r |> 
  ggplot(aes(
    x=Year,
    y=prop_na,
    col=Region))+
  geom_line(position = position_jitter()) +
  theme_bw() +
  ggtitle('% NAs in Electricity Statistics per Year and Region') +
  scale_y_continuous('% NAs',labels=scales::percent_format()) +
  scale_color_discrete(breaks=~.x[!is.na(.x)]) +
  theme(legend.position="bottom")
```
In the first years, Middle East is the region with most NAs (<40%). Around 2014 we see an improvement for this region. We add jitter to observe the pattern for regions with least share of missing values. In particular Asia, Europe, Eurasia and North America have small shares of NAs (<5%).

Again, we assess some type of issue with data on 2023 due to an increase of shares of missing values.

### Electricity statistics (MW/GWh) by Country/area, Technology, Data Type, Grid connection and Year

```{r}
#| message: false
# read data
df_electstats_country <- read_csv(
  "./raw_data/irena_electstats_country.csv"
)

attr(df_electstats_country, 'spec') <- NULL
```


```{r}
# electricity statistics column is saved as character. 
# in particular, some rows have value equal "-", we label them as NA 
# convert electricity statistics to numeric
df_electstats_country <- df_electstats_country |> 
  mutate(
    across(
      starts_with('Elec'),
      ~ifelse(.x=='-',NA,.x)
    )
  ) |> 
  mutate(
    across(
      starts_with('Elec'),
      as.numeric
      )
  )
```

Let's review the proportion of NAs by column:

```{r}
# review proportion of NAs per column: global
df_electstats_country |> 
  summarise(
    across(
      everything(),
      ~sum(ifelse(is.na(.x),1,0))
    ),
    n=n()
  ) |> 
  pivot_longer(
    cols=-n,
    names_to = 'variable',
    values_to = 'n_na'
  ) |> 
  mutate(prop_na=n_na/n) |> 
  data.table()
```
Only Electricity statistics column presents NAs, around 75% of the total observations.

Let's review the pattern by Year

```{r}
# review proportion of NAs by Year
df_electstats_country_year <- df_electstats_country |> 
  group_by(Year) |> 
  summarise(
    n = n(),
    n_country = n_distinct(`Country/area`),
    across(
      starts_with('Elect'),
      ~sum(ifelse(is.na(.x),1,0))
    )
  ) |>
  rename(n_na=starts_with('Elect')) |> 
  mutate(prop_na=n_na/n)

# graph share of NAs in Electric Statistics per Year
df_electstats_country_year |> 
  ggplot(aes(x=Year,y=prop_na)) +
  geom_line() +
  theme_bw() +
  ggtitle('% NAs in Electricity Statistics per Year') +
  scale_y_continuous('% NAs',labels=scales::percent_format())

```

The share of missing values present a downtrend but on year 2023 there's a sudden jump up to 85%.

Let's assess whether % of NAs depend on which Technology is measured

```{r}
#| fig-height: 7

# review proportion of NAs per by Year and Technology
df_electstats_country_year_tec <- df_electstats_country |> 
  group_by(Year,Technology) |> 
  summarise(
    n = n(),
    across(
      starts_with('Elect'),
      ~sum(ifelse(is.na(.x),1,0))
    )
  ) |>
  rename(n_na=starts_with('Elect')) |> 
  mutate(prop_na=n_na/n)

#  % share of NAs per by Year and Technology
df_electstats_country_year_tec |> 
  mutate(Year=factor(Year)) |> 
  ggplot(aes(x=Technology,y=Year,fill=prop_na)) +
  geom_tile() +
  scale_fill_gradient2(
    "% of NAs",
    low='green',
    mid='white',
    high='orange',
    midpoint = 0.5,
    labels=scales::percent_format()
  ) +
  theme_bw() +
  ggtitle(paste0(
  '% NAs in Electricity Statistics',
  ' per Year and Technology')
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

For most Technologies we have big shares of missing values (>50%). For Solar energy we the share decreases across time. Just for Oil and aggregated values such as Total Non-Renewable and Total Renewable the share of NAs is below 50% consistently.

Let's analyze for grid connection

```{r}
#| fig-height: 7

# review proportion of NAs per by Year and Grid connection
df_electstats_country_year_grid <- df_electstats_country |> 
  group_by(Year,`Grid connection`) |> 
  summarise(
    n = n(),
    across(
      starts_with('Elect'),
      ~sum(ifelse(is.na(.x),1,0))
    )
  ) |>
  rename(n_na=starts_with('Elect')) |> 
  mutate(prop_na=n_na/n) 

#  % share of NAs per by Year and Grid connection
df_electstats_country_year_grid |> 
  mutate(Year=factor(Year)) |> 
  ggplot(aes(x=`Grid connection`,y=Year,fill=prop_na)) +
  geom_tile() +
  scale_fill_gradient2(
    "% of NAs",
    low='green',
    mid='white',
    high='orange',
    midpoint = 0.5,
    labels=scales::percent_format()
  ) +
  theme_bw() +
  ggtitle(paste0(
    '% NAs in Electricity Statistics',
    ' per Year and Grid Connection')
    ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

Regardless of the grid connection the share of missing value is high (>60%)

For which countries do we have most data?

```{r}
#| fig-height: 7

# review proportion of NAs per by Year and Country
df_electstats_country_year_country <- df_electstats_country |> 
  group_by(Year,`Country/area`) |> 
  summarise(
    n = n(),
    across(
      starts_with('Elect'),
      ~sum(ifelse(is.na(.x),1,0))
    )
  ) |>
  rename(n_na=starts_with('Elect')) |> 
  mutate(prop_na=n_na/n) |> 
  mutate(Country=ifelse(prop_na<0.5,`Country/area`,NA)) |> 
  mutate(Country = str_remove_all(Country,'\\(the\\)'))  |> 
  mutate(Country = case_when(
    str_detect(Country,'United Kingdom') ~ 'UK',
    str_detect(Country,'United States') ~ 'US',
    str_detect(Country,'Korea') ~ 'South Korea',
    str_detect(Country,'Viet Nam') ~ 'Vietnam',
    TRUE ~ Country
  ))

# graph % NA per Year and Country
# color countries with %NAs < 50%
df_electstats_country_year_country |> 
  ggplot(aes(
    x=Year,
    y=prop_na,
    group=`Country/area`,
    col=Country))+
  geom_line(position=position_jitter()) +
  theme_bw() +
  ggtitle('% NAs in Electricity Statistics per Year and Country') +
  scale_y_continuous('% NAs',labels=scales::percent_format()) +
  scale_color_discrete(breaks=~.x[!is.na(.x)]) +
  theme(legend.position="bottom")
```

Highlighting countries which share of NAs is less than 50% only 12 met this criteria out of 200+ countries. 

### Renewable energy share of electricity capacity and generation (%) by Region/country/area, Indicator and Year

```{r}
#| message: false
# read data
df_share <- read_csv(
  "./raw_data/irena_share.csv"
)

attr(df_share, 'spec') <- NULL 
```


```{r}
# column is saved as character. 
# in particular, some rows have value equal "-", we label them as NA 
# convert renewable column to numeric
# divide pct by 100
# includes region and country/area.
# Let's split data an analyze separately
df_share <- df_share|> 
  mutate(
    across(
      starts_with('Renewable'),
      ~ifelse(.x=='-',NA,.x)
    )
  ) |> 
  mutate(
    across(
      starts_with('Renewable'),
      as.numeric
    )
  ) |> 
  mutate(
    across(
      starts_with('Renewable'),
      ~.x/100
    )
  )

# has countries and regions together split 
c_names <- df_share$`Region/country/area` |> 
  unique()

# by region
df_share_region <- df_share |> 
  rename(Region=`Region/country/area`) |> 
  filter(Region %in% c_names[1:10])

# review proportion of NAs per column: global
df_share_region |> 
  summarise(
    across(
      everything(),
      ~sum(ifelse(is.na(.x),1,0))
    ),
    n=n()
  ) |> 
  pivot_longer(
    cols=-n,
    names_to = 'variable',
    values_to = 'n_na'
  ) |> 
  mutate(prop_na=n_na/n) |> 
  data.table()
```

In this case just 2.3% of data is missing. Let's review the pattern if we group by Year

```{r}
#| fig-height: 5
# review proportion of NAs by Year
df_share_region_year <- df_share_region |> 
  group_by(Year) |> 
  summarise(
    n = n(),
    n_country = n_distinct(Region),
    across(
      starts_with('Renewable'),
      ~sum(ifelse(is.na(.x),1,0))
    )
  ) |>
  rename(n_na=starts_with('Renewable')) |> 
  mutate(prop_na=n_na/n)

# graph share of NAs in Renewable per Year
df_share_region_year |> 
  ggplot(aes(x=Year,y=prop_na)) +
  geom_line() +
  theme_bw() +
  ggtitle(paste0(
    '% NAs in Renewable energy share of',
    ' electricity capacity \nand generation per Year')) +
  scale_y_continuous('% NAs',labels=scales::percent_format())
```

In this case this 2.3% is explained due to 2023 not having any value for this dataset. Hence if we exclude this year our data at a regional level has no missing values.

Let's assess our data at a country-level

```{r}
# at position 11 starts countries
df_share_country <- df_share |> 
  rename(`Country/area`=`Region/country/area`) |> 
  filter(`Country/area` %in% c_names[11:length(c_names)])

# review proportion of NAs per column: global
df_share_country |> 
  summarise(
    across(
      everything(),
      ~sum(ifelse(is.na(.x),1,0))
    ),
    n=n()
  ) |> 
  pivot_longer(
    cols=-n,
    names_to = 'variable',
    values_to = 'n_na'
  ) |> 
  mutate(prop_na=n_na/n) |> 
  data.table()
```

In this case, the data set have ~10% of missing values

Again, let's review whether this is explained just by Year

```{r}
#| fig-height: 5

# review proportion of NAs group by Year
df_share_country_year <- df_share_country |> 
  group_by(Year) |> 
  summarise(
    n = n(),
    n_country = n_distinct(`Country/area`),
    across(
      starts_with('Renewable'),
      ~sum(ifelse(is.na(.x),1,0))
    )
  ) |>
  rename(n_na=starts_with('Renewable')) |> 
  mutate(prop_na=n_na/n)

# graph share of NAs in Electric Statistics per Year
df_share_country_year |> 
  ggplot(aes(x=Year,y=prop_na)) +
  geom_line() +
  theme_bw() +
  ggtitle(paste0(
    '% NAs in Renewable energy share of',
    ' electricity capacity \nand generation per Year')) +
  scale_y_continuous('% NAs',labels=scales::percent_format())
```

In this case there are missing values in different years with a downtrend and a spike on 2023.

Let's assess how missing values are with Year and Indicator column

```{r}
#| fig-height: 7

# review proportion of NAs per by Year and Indicator
df_share_country_year_ind <- df_share_country |> 
  group_by(Year,Indicator) |> 
  summarise(
    n = n(),
    across(
      starts_with('Renewable'),
      ~sum(ifelse(is.na(.x),1,0))
    )
  ) |>
  rename(n_na=starts_with('Renewable')) |> 
  mutate(prop_na=n_na/n)

#  % share of NAs per by Year and Indicator
df_share_country_year_ind |> 
  mutate(Year=factor(Year)) |> 
  ggplot(aes(x=Indicator,y=Year,fill=prop_na)) +
  geom_tile() +
  scale_fill_gradient2(
    "% of NAs",
    low='green',
    mid='white',
    high='orange',
    midpoint = 0.5,
    labels=scales::percent_format()
  ) +
  theme_bw() +
  ggtitle(paste0(
    '% NAs in Renewable energy share of electricity capacity',
    ' \nand generation per Year and Indicator')) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Aside from 2023, missing value share for Indicator is below ~25% across time.

After seeing this we would expect a lot of countries to have values in the metric of interest. Let's visualize how the share of countries that have less than a certain threshold of missing values.

```{r}
#| fig-height: 7

# review proportion of NAs per by Year and Country
df_share_country_year_c <- df_share_country |> 
  group_by(Year,`Country/area`) |> 
  summarise(
    n = n(),
    across(
      starts_with('Renewable'),
      ~sum(ifelse(is.na(.x),1,0))
    )
  ) |>
  rename(n_na=starts_with('Renewable')) |> 
  mutate(prop_na=n_na/n) |> 
  mutate(flag=ifelse(prop_na<0.5,1,0)) |> 
  group_by(Year) |> 
  summarise(less_50_na=sum(flag)/n())

# share of countries that meet criteria
df_share_country_year_c  |> 
  ggplot(aes(
    x=Year,
    y=less_50_na))+
  geom_line() +
  theme_bw() +
  ggtitle(paste0(
    '% of countries with less than 50% of NAs',
    ' in Renewable energy share of \nelectricity capacity',
    ' and generation per Year')
    ) +
  scale_y_continuous('% NAs',labels=scales::percent_format()) 
```

For this particular dataset from the 200+ countries above 75% of them have less than 50% of missing values. Unlike the first dataset we can do an analysis at region as well at country-level.

### Public Investments (2021 million USD) by Country/area, Technology and Year

```{r}
#| message: false
# read data
df_finance <- read_csv(
  "./raw_data/irena_finance.csv"
)

attr(df_finance, 'spec') <- NULL
```

Let's analyze whether this dataset has missing values or not.

```{r}
# review proportion of NAs per column: global
df_finance |> 
  summarise(
    across(
      everything(),
      ~sum(ifelse(is.na(.x),1,0))
    ),
    n=n()
  ) |> 
  pivot_longer(
    cols=-n,
    names_to = 'variable',
    values_to = 'n_na'
  ) |> 
  mutate(prop_na=n_na/n) |> 
  data.table()
```

For finance data it looks like no missing values are reported.
